---
name: Dr. Joe Onoufriou
qualifications: BSc, MRes, PhD
position: "Post-Doctoral Researcher"
address: "Scottish Association for Marine Science"
phone: +44 (0)7921 389727
email: "jojo.onoufriou@gmail.com"
twitter: JojoOnoufriou
github: JojoOno
orcid: 0000-0001-5509-9667
date: "`r format(Sys.time(), '%B %Y')`"
headcolor: 009ACD
output: 
  vitae::awesomecv:
    keep_tex: false
header-includes:
  \usepackage{float}
  \usepackage{multicol}
  \usepackage{colortbl}
  \arrayrulecolor{white}
  \usepackage{hhline}
  \definecolor{light-gray}{gray}{0.95}

---

```{r setup, include=FALSE}
library(vitae)
library(tidyverse)
library(fontawesome)
library(emojifont)
library(kableExtra)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

\faIcon{lightbulb} Research Interests
============================

An applied ecologist with extensive field and analytical experience, particularly in predator foraging ecology, oceanographiy and effects of renewable energy installations on population distributions and individual fitness. Experienced in academia and collaborating with industry and policy stakeholders, resulting in a high level of competency producing science which is robust, defensible and communicable to a wide audience. 

Key Words: Spatial and movement statistics; applied ecology; marine mammal ecology and pathology


\faIcon{graduation-cap} Education
============================

```{r}
library(tibble)
tribble(
  ~ degree, ~ uni, ~loc, ~ dates, ~ details,
  "PhD", "Univeristy of St Andrews", "St Andrews, Scotland", "Nov. 2015 - Jan. 2020",
   "Thesis title: 'Movement Ecology of seals in a tidally energetic environment: Implications for interactions with tidal energy devices'",
  "MRes", "University of St Andrews", "St Andrews, Scotland", "Sep. 2011 - Aug. 2012",  
    "Thesis title: ‘A validated step-wise approach to drift-dive classification using high resolution accelerometery data’",
  "BSc (Hons)", "University of Southampton", "Southampton, England", "Sep. 2006 - Sep. 2010",
   "Marine Biology and Oceanography. Thesis title: 'The effects of the toxic dinoflagellate, \\textit{Coolia monotis}, on respiration rate and survival of brine shrimp'") %>% 
  detailed_entries(degree, dates, uni, loc, details, .protect=FALSE)
```

\faIcon{university} Relevant Employment
============================
```{r}
tribble(
  ~ role, ~ company, ~ loc, ~ dates, ~ details,
  "Post-doctoral researcher", "University of the Highlands and Islands", "Oban, Scotland", "Apr. 2020 - Present", 
    list("Investigating the effects of tidal turbine sound exposure on seal movement and distribution.",
         "Providing analytical, administrative and project focussed support and mentoring to PhD students.",
         "Using animal movement data to help refine estimates of collision risk between marine mammals and tidal turbines."),
   "Tutor and lecturer", "University of St Andrews", "St Andrews, Scotland", "Sep. 2019 - present", 
    list("Guest lecturer in marine animal movement and spatial analysis (2 lectures per semester) - BL5110", "Developing and leading lab practicals on the use of R to quantify movement behaviour using telemetry data - BL5122", "Demonstrating for lab practicals on quantitative methods in ecology and GIS for ecologists - BL5110"),
  "Post-doctoral researcher", "University of St Andrews", "St Andrews, Scotland", "Jan. 2020 - Apr. 2020", 
    list("Supporting delivery of the Marine Scotland Scientific Support Programme through dissemination of spatial analysis to Scottish ministers",
         "Providing consultation on collision risk and impact of operational tidal turbine arrays"),
  "Post-graduate research assistant", "University of St Andrews", "St Andrews, Scotland", "Sep. 2014 - Nov. 2016", 
    list("Quantifying the spatial overlap between seals and shipping traffic in the Moray Firth, Scotland.",
         "Using telemetry data and AIS monitoring to predict seal and shipping distribution across the study site, scaling the seal predictions with aerial survey count data at individual haul-outs and estimating hot-spots of interest where seal-shipping overlap appeared to present a potential risk to local populations."),
   "Post-graduate research assistant", "University of St Andrews", "St Andrews, Scotland", "Apr. 2013 - Nov. 2014", 
    list("Investigating unusual mortality events (“Corkscrew Lesions”) in seals around the UK.",
         "Curating databases of seal stranding cases around the Scottish coast.", "Designing and conducting experiments to test the hypothetical link between corkscrew lesions and ship propellers.", "Leading the investigation into grey seals as a potential source of unusual mortality events in seals using pathological analyses, telemetry data and recorded behavioural data"
         )
  # "Data mining intern", "iSelect", "Cheltenham, Australia", "Feb. 2015 - Mar. 2015", "Improved business reporting with interactive visualisations, and model-based anomaly detection.",
) %>% 
  detailed_entries(role, dates, company, loc, details)
```

\faIcon{book} Selected Publications
============================

For a complete list of publications and reports see [*my Google Scholar profile (link)*](https://scholar.google.com/citations?hl=en&user=bh4T-04AAAAJ).

```{r pubs}
tribble(
  ~ name, ~ journal, ~ date, ~type, ~ bib,
    "Foraging in dynamic habitats: the importance of considering flow in animal movement analyses", "Functional Ecology", "~2021", "Journal Article", "\\textbf{Onoufriou, J.}, Russell, D.J.F., Thompson, D., Moss, S., O'Hara Muraye, R.,        Hastie, G. Functional Ecology. https://doi.org/10.1016/j.renene.2021.08.052",
  
    "Quantifying the effects of tidal turbine array operations on the distribution of marine mammals: implications for collision risk and spatial planning", "Renewable Energy", "Aug. 2021", "Journal Article", 
   "\\textbf{Onoufriou, J.}, Russell, D.J.F., Thompson, D., Moss, S., Hastie, G. Renewable Energy. 2021; 180: 157-165y.",
  
    "Empirical determination of severe trauma in seals from collisions with tidal turbine blades", "Journal of Applied Ecology", "Jan. 2019", "Journal Article", 
  "\\textbf{Onoufriou, J.}, Brownlow, A, Moss, S, Hastie, G, Thompson, D. Empirical determination of severe trauma in seals from collisions with tidal turbine blades. J Appl Ecol. 2019; 56: 1712– 1724. https://doi.org/10.1111/1365-2664.13388",
  
     "Seals and shipping: quantifying population risk and individual exposure to vessel noise", "Journal of Applied Ecology", "Apr. 2017", "Journal Article", 
  "Jones, E.L., Hastie, G.D., Smout, S., \\textbf{Onoufriou, J.}, Merchant, N.D., Brookes, K.L. and Thompson, D. Seals and shipping: quantifying population risk and individual exposure to vessel noise. J Appl Ecol, 54: 1930-1940. https://doi.org/10.1111/1365-2664.12911",
  
   "Corkscrew Seals: Grey Seal (\\textit{Halichoerus grypus}) infanticide and cannibalism may indicate the cause of spiral lacerations in seals", "PLOS ONE", "Jun. 2016", "Journal Article", 
  "Brownlow A, \\textbf{Onoufriou, J.}, Bishop A, Davison N, Thompson D. Corkscrew Seals: Grey Seal (\\textit{Halichoerus grypus}) Infanticide and Cannibalism May Indicate the Cause of Spiral Lacerations in Seals. PLOS ONE 11(6): e0156464. https://doi.org/10.1371/journal.pone.0156464",

   "Cannibalism by a male grey seal (\\textit{Halichoerus grypus}) in the North Sea", "Aquatic Mammals", "Jun. 2016", "Journal Article", 
  "Bishop, A. M., \\textbf{Onoufriou, J.}, Moss, S, Pomeroy, P, Twiss, S.D. Cannibalism by a male grey seal (\\textit{Halichoerus grypus}) in the North Sea. Aquatic Mammals: Vol. 42, Iss. 2. 137-143. DOI:10.1578/AM.42.2.2016.137",
  
  "Investigations into the interactions between harbour seals (\\textit{Phoca vitulina}) and vessels in the inner Moray Firth", "Scottish Marine and Freshwater Science", "Mar. 2016", "Government Report", 
  "\\textbf{Onoufriou, J.}, Jones, E., Hastie, G. Thompson, D. Investigations into the interactions between harbour seals (Phoca vitulina) and vessels in the inner Moray Firth. Scottish marine and freshwater science: 7(15). DOI: 10.7489/1805-1",
  
  "Testing the hypothetical link between shipping and unexplained seal deaths", "Marine Scotland Scientific Support Reserach Programme", "Mar. 2016", "Government Report", 
  "\\textbf{Onoufriou, J.}, Thompson, D., Brownlow, A. Testing the hypothetical link between shipping and unexplained seal deaths. MMS/001/11-USD2. http://www.smru.st-and.ac.uk/documents/1926.pdf"
  
  ) %>% 
 
  detailed_entries(with = name, what = journal, when = date, where = type, why = bib, .protect=FALSE)

```


\faIcon{wrench} Skills
================================

\faIcon{laptop} Software and Programming  
---------------------------------

```{r}
tribble(
   ~`Coding Languages`, ~Software, ~`Data Types`, ~Other,
  "R -- SQL -- MatLab",
  "QGIS -- ArcGIS -- Manifold -- GIMP -- Zotero -- MSOffice",
   "GPS -- ARGOS -- TDR -- AIS -- Accelerometry",
  "Git -- Markdown -- LaTex"
 
) %>%
  kable(booktabs = T, align = 'c', escape = F, format = "latex") %>%
  kable_styling(full_width = F, font_size = 9, latex_options="hold_position") %>% 
  row_spec(0, bold = T, color = "#009acd") %>% 
  row_spec(1, bold = F, color = "#7f7f7f") %>% 
  column_spec(column = 1:4, width = '3.75cm')
```

\faIcon{chart-bar} Statistics
---------------------------------

Linear regression and extensions including mixed effects models, generalised additive models (GAM) and generalised estimating equations (GEE) applied to species distribution, habitat preference and animal behaviour analyses.

State-space models and hidden Markov models (HMMs) as tools for inferring animal behaviour in discrete and continous time including assessing biotic and abiotic covariate effects on state transitions and discrete behaviours, and establishing environmental drivers of movement to generate informative data-streams.\  
  
&nbsp; 

\faIcon{leaf} Fieldwork
---------------------------------

8 years experience of capture, handling and tag attachment of phocid seals (grey and harbour seals).

3 years experience as an aerial survey technician for breeding and moulting surveys of UK phocid seals (grey and harbour seals).

2 field seasons assisting capture and handling of seabirds (puffins and shags) on a breeding colony.

Experienced small-boat operator - Royal Yacht Association (RYA) Power Boat level 2

Marine and Coastguard Agency (MCA) approved sea-survival


\faIcon{comments} Conference presentations and workshops
================================

```{r pres}
tribble(
  ~ name, ~ title, ~ conference, ~year, ~ presentation,
  
  "Onoufriou, J., Russell, D.J.F, Thompson, D. Moss, S.E. & Hastie, G.", "Variable diving behaviour of seals in tidally energetic channels", "MASTS: Marine Alliance for Science and Technology forum. Online.", "2021", "Oral Presentation",
  
     "Onoufriou, J., Russell, D.J.F, Thompson, D. Moss, S.E. & Hastie, G.", "Quantifying the effects of tidal turbine array operations on the distribution of marine mammals: implications for collision risk ", "EIMR: Environmental Interactions of Marine Renewable Energy Technologies. Online.", "2020", "Oral Presentation",
  
   "Onoufriou, J., Russell, D.J.F, Thompson, D. Moss, S.E. & Hastie, G.", "Overt avoidance behaviour of seals in response to tidal turbine array operations", "ScotMER symposium, Marine Scotland. Online.", "2020", "Oral Presentation",
  
  "Onoufriou, J., Russel, D.J.F., Thompson, D., O’Hara Murray, R. Moss, S. & Hastie, G.", "Foraging in dynamic habitats: the importance of considering flow in animal movement analyses", "World Marine Mammal Conference. Barcelona, Spain.", "2019", "Speed talk",
  
  "Onoufriou, J., Russel, D.J.F., Thompson, D., O’Hara Murray, R. Moss, S. & Hastie, G.", "Foraging plasticity and extrinsic drivers of activity budgets in a tidally energetic system.", "SEECC 2019: Scottish Ecology, Environment and Conservation Conference. Glasgow, UK.", "2019", "Invited Talk",
  
  "Onoufriou, J., Brownlow, A., Moss, S., Hastie, G. & Thompson, D.", "Empirical determination of severe trauma in seals from collisions with tidal turbine blades.", "ScotMER symposium, Marine Scotland.", "2019", "Invited Talk",
  
  "Onoufriou, J., Brownlow, A., Moss, S., Hastie, G. & Thompson, D.", "Empirical determination of severe trauma in seals from collisions with tidal turbine blades.", "SIMC: Scotland’s International Marine Conference.", "2019", "Poster",
  
   "Onoufriou, J., Thompson, D., Brownlow, A., Sparling, C. & Hastie, G.", "Experimental determination of a mortality threshold for collisions between marine megafauna and tidal turbines.", "EIMR: Environmental Interactions of Marine Renewable Energy Technologies.", "2018", "Oral Presentation",
  
  "Onoufriou, J., Clair Evers, Dave Thompson, Andrew Brownlow & Gordon Hastie.", "Seal collision trials and fine-scale tracking around an operating turbine array.", "Workshop on environmental impacts of tidal energy industry: 22nd Biennial Conference for the Society of Marine Mammalogy. Halifax, NS", "2017", "Invited Talk",
  
   "Onoufriou, J., Brownlow, A., Moss, S., Hastie, G. & Thompson, D.", "Corkscrew seals: grey seal infanticide and cannabilism may indicate the cause of spiral lacerations in seals", "21st Biennial Conference for the Society of Marine Mammalogy. San Francisco, CA", "2015", "Workshop coordinator",
  
  "Onoufriou, J., Brownlow, A & Thompson, D. ", "Unravelling the mystery of corkscrew seals.", "UKIRSCSMM: UK and Ireland Student Chapter of the Society of Marine Mammalogy Conference. Bangor, UK", "2015", "Workshop coordinator",
  
  "Onoufriou, J., Fedak, M. & Thompson, D.", "A validated, step-wise approach to drift-dive classification using high resolution accelerometery", "20th Biennial Conference for the Society of Marine Mammalogy. Dunedin, NZ", "2013", "Oral Presentation"
  
 
  ) %>% 
 
  detailed_entries(with = title, what = name, when = year, where = presentation, why = conference)

```

\newpage

\faIcon{users} Students
================================

Supervised/advised 
---------------------------------
2020-21 Charlotte Findlay, PhD (co-supervised)  
2019-20 Tierney Carter, MSc (co-supervised)  
2018-19 Laura Palmer, MSc (co-supervised)  
2017-18, Claire Evers, MSc (co-supervised)  

PhD students mentored under The Bryden Centre Doctoral Training Program 2020-2021
---------------------------------
Natalie Issakson (Environmental Research Institute), Inne Withouck (NAFC Marine Centre), Morag Cooper (Inverness College), Rowland Fraser (Inverness College), Monika Kosecka (Scottish Association for Marine Science)

\faIcon{suitcase} Professional service
================================

Reviewer for Scientific Journals
---------------------------------
International Marine Energy Journal, Scientific Reports, Aquatic Conservation, PLOS One, Journal of Marine Science and Engineering 

Editorial positions
---------------------------------
Review editor: Frontiers in Physiology (physio-logging)